### 2018-06-27 pubs/itp-2018/sections/proof-reconstruction/canonicalize.tex
First attempt to remove complexity measure in sec4

The \canonicalize rule is mainly used by \Metis to introduce the negation of a
subgoal in its refutation proof. When the input formula is an axiom, definition
or hypothesis introduced in the \TPTP file of the problem, the \canonicalize
rule returns its negative normal form. In the remaining cases, as far as we
know, \Metis uses the \canonicalize rule to performs a sort of simplifications
to remove the tautologies listed in~Fig.~\ref{fig:redundancies} in the
conjunctive normal form of the input formula.

@@

\begin{empheq}[box=\fcolorbox{bocolor}{bgcolor}]{equation}
  \label{eq:nnf}
  \begin{aligned}
    &\hspace{.495mm}\fnnf_{1} : \Prop \to \Nat \to \Prop\\
    &\begin{array}{llll}
    \fnnf_{1} &(φ₁ ∧ φ₂)     &(\suc~n) &= \fcanon_{∧}~(\fassoc_{∧}~(\fnnf_{1}~φ₁~n ∧ \fnnf_{1}~φ₂~n))\\
    \fnnf_{1} &(φ₁ ∨ φ₂)     &(\suc~n) &= \fcanon_{∨}~(\fassoc_{∨}~(\fnnf_{1}~φ₁~n ∨ \fnnf_{1}~φ₂~n))\\
    \fnnf_{1} &(φ₁ ⇒ φ₂)     &(\suc~n) &= \fcanon_{∨}~(\fassoc_{∨}~(\fnnf_{1}~((¬ φ₁) ∨ φ₂)~n))\\
    \fnnf_{1} &(¬ (φ₁ ∧ φ₂)) &(\suc~n) &= \fcanon_{∨}~(\fassoc_{∨}~(\fnnf_{1}~((¬ φ₁) ∨ (¬ φ₂))~n))\\
    \fnnf_{1} &(¬ (φ₁ ∨ φ₂)) &(\suc~n) &= \fcanon_{∧}~(\fassoc_{∧}~(\fnnf_{1}~((¬ φ₁) ∧ (¬ φ₂))~n))\\
    \fnnf_{1} &(¬ (φ₁ ⇒ φ₂)) &(\suc~n) &= \fcanon_{∧}~(\fassoc_{∧}~(\fnnf_{1}~((¬ φ₂) ∧ φ₁)~n))\\
    \fnnf_{1} &(¬ (¬ φ))     &(\suc~n) &= \fnnf_{1}~φ₁~n\\
    \fnnf_{1} &(¬ ⊤)         &(\suc~n) &= ⊥\\
    \fnnf_{1} &(¬ ⊥)         &(\suc~n) &= ⊤\\
    \fnnf_{1} &φ             &\zero    &= φ
    \end{array}
  \end{aligned}
\end{empheq}

The function $\fnnf$ in~\eqref{eq:fnnf} is defined similarly as the \fstrip
function in~\eqref{eq:strip}, the complexity measure function $\fnnf_{cm}$ is
defined in Appendix in~\cite{Prieto-Cubides2017a}.

>>>>

The \canonicalize rule is mainly used by \Metis to introduce the negation of a
subgoal in its refutation proof. When the input formula is an axiom, definition
or hypothesis introduced in the \TPTP file of the problem, the \canonicalize
rule returns its negative normal form. In the remaining cases, as far as we
know, \Metis uses the \canonicalize rule to perform a sort of simplifications
to remove the tautologies listed in~Fig.~\ref{fig:redundancies} from the
conjunctive normal form of the input formula.

@@

\begin{empheq}[box=\fcolorbox{bocolor}{bgcolor}]{equation}
  \label{eq:nnf}
  \begin{aligned}
    &\hspace{.495mm}\fnnf_{1} : \Prop \to \Bound \to \Prop\\
    &\begin{array}{llll}
    \fnnf_{1} &(φ₁ ∧ φ₂)     &(\suc~n) &= \fcanon_{∧}~(\fassoc_{∧}~(\fnnf_{1}~φ₁~n ∧ \fnnf_{1}~φ₂~n))\\
    \fnnf_{1} &(φ₁ ∨ φ₂)     &(\suc~n) &= \fcanon_{∨}~(\fassoc_{∨}~(\fnnf_{1}~φ₁~n ∨ \fnnf_{1}~φ₂~n))\\
    \fnnf_{1} &(φ₁ ⇒ φ₂)     &(\suc~n) &= \fcanon_{∨}~(\fassoc_{∨}~(\fnnf_{1}~((¬ φ₁) ∨ φ₂)~n))\\
    \fnnf_{1} &(¬ (φ₁ ∧ φ₂)) &(\suc~n) &= \fcanon_{∨}~(\fassoc_{∨}~(\fnnf_{1}~((¬ φ₁) ∨ (¬ φ₂))~n))\\
    \fnnf_{1} &(¬ (φ₁ ∨ φ₂)) &(\suc~n) &= \fcanon_{∧}~(\fassoc_{∧}~(\fnnf_{1}~((¬ φ₁) ∧ (¬ φ₂))~n))\\
    \fnnf_{1} &(¬ (φ₁ ⇒ φ₂)) &(\suc~n) &= \fcanon_{∧}~(\fassoc_{∧}~(\fnnf_{1}~((¬ φ₂) ∧ φ₁)~n))\\
    \fnnf_{1} &(¬ (¬ φ))     &(\suc~n) &= \fnnf_{1}~φ₁~n\\
    \fnnf_{1} &(¬ ⊤)         &(\suc~n) &= ⊥\\
    \fnnf_{1} &(¬ ⊥)         &(\suc~n) &= ⊤\\
    \fnnf_{1} &φ             &\zero    &= φ
    \end{array}
  \end{aligned}
\end{empheq}

The function $\fnnf$ in~\eqref{eq:fnnf} is defined similarly as the \fstrip
function in~\eqref{eq:strip} with a function $\fnnf_{cm}$ that finds a bound
for its recursive calls (See its definition in Appendix in~\cite{Prieto-Cubides2017a}).

### 2018-06-27 pubs/itp-2018/sections/proof-reconstruction/strip.tex
First attempt to remove complexity measure in sec4

\begin{mainlemma}
  \label{lem:inv-uh-lem}
Let be $\text{n} : \Nat$. If $Γ ⊢ \fuh₁~φ~n$ then $Γ ⊢ φ$ where

\begin{empheq}[box=\fcolorbox{bocolor}{bgcolor}]{equation}
  \label{eq:uh-structured}
  \begin{aligned}
  &\hspace{.495mm} \fuh_{1} : \Prop → \Nat → \Prop\\
  &\begin{array}{llll}
  \fuh_{1} &(φ₁ ⇒ (φ₂ ⇒ φ₃)) &(\suc~n) &= \fuh_{1}~((φ₁ ∧ φ₂) ⇒ φ₃)~n\\
  \fuh_{1} &(φ₁ ⇒ (φ₂ ∧ φ₃)) &(\suc~n) &= \fuh_{1}~(φ₁ ⇒ φ₂)~n ∧ \fuh_{1}~(φ₁ ⇒ φ₃)~n\\
  \fuh_{1} &φ &n &= φ.
  \end{array}
  \end{aligned}
\end{empheq}
\end{mainlemma}

The second argument in the $\fuh_1$ function is a natural number that constrains
the number of recursive calls. To provide such a number given a formula
$\varphi$, we call the $\fuh_{cm}$ function defined in~\eqref{eq:uh-complexity}.
This function finds what we call the \emph{complexity measure} of the function
$\fuh_1$. For the convenience in the following descriptions, we have the
function $\fuh : \Prop → \Prop$ for $\varphi \mapsto \fuh_{1}~φ~(\fuh_{cm}~φ)$.
\begin{empheq}[box=\fcolorbox{bocolor}{bgcolor}]{equation}
  \label{eq:uh-complexity}
  \begin{aligned}
    &\hspace{.495mm}\fuh_{cm} : \Prop → \Nat\\
    &\begin{array}{llll}
    \fuh_{cm}~(φ₁ ⇒ (φ₂ ⇒ φ₃)) &= \fuh_{cm}~φ₃ + 2\\
    \fuh_{cm}~(φ₁ ⇒ (φ₂ ∧ φ₃)) &= \fmax~(\fuh_{cm}~φ₂)~(\fuh_{cm}~φ₃) + 1\\
    \fuh_{cm}~φ                &= 0.
    \end{array}
  \end{aligned}
\end{empheq}

Now, we define the $\fstrip$ function in \eqref{eq:strip} in a similar way as we
did above for the $\fuh$ function. We have a definition of the $\fstrip₁$ function
in~\eqref{eq:strip-fixed} and a complexity measure denoted by $\fstrip_{cm}$
and defined in~\cite{Prieto-Cubides2017a}.
% \vspace*{-5mm}
\begin{empheq}[box=\fcolorbox{bocolor}{bgcolor}]{align}
  \label{eq:strip}
  \begin{split}
  &\fstrip : \Prop → \Prop\\
  &\fstrip~φ~ = \fstrip₁~φ~(\fstrip_{cm}~φ).
  \end{split}
\end{empheq}

The $\fstrip$ function yields the conjunction of subgoals that implies the goal
of the problem in the \Metis \TSTP derivations. Its definition comes mainly from
the reading of the \Metis source code and its \TSTP derivations.

\begin{empheq}[box=\fcolorbox{bocolor}{bgcolor}]{equation}
  \label{eq:strip-fixed}
  \begin{aligned}
  &\hspace{.495mm}\fstrip₁ : \Prop → \Nat → \Prop\\
  &\begin{array}{llll}
  \fstrip₁ &(φ₁ ∧ φ₂)     &(\suc~n) &= \fuh~(\fstrip₁~φ₁~n) ∧ \fuh~(φ₁ ⇒ \fstrip₁~φ₂~n)\\
  \fstrip₁ &(φ₁ ∨ φ₂)     &(\suc~n) &= \fuh~((¬ φ₁) ⇒ \fstrip₁~φ₂~n)\\
  \fstrip₁ &(φ₁ ⇒ φ₂)     &(\suc~n) &= \fuh~(φ₁ ⇒ \fstrip₁~φ₂~n)\\
  \fstrip₁ &(¬ (φ₁ ∧ φ₂)) &(\suc~n) &= \fuh~(φ₁ ⇒ \fstrip₁~(¬ φ₂)~n)\\
  \fstrip₁ &(¬ (φ₁ ∨ φ₂)) &(\suc~n) &= \fuh~(\fstrip₁~(¬ φ₁)~n) ∧ \fuh~((¬ φ₁) ⇒ \fstrip₁~(¬ φ₂)~n)\\
  \fstrip₁ &(¬ (φ₁ ⇒ φ₂)) &(\suc~n) &= \fuh~(\fstrip₁~φ₁~n) ∧ \fuh~(φ₁ ⇒ \fstrip₁~(¬ φ₂)~n)\\
  \fstrip₁ &(¬ (¬ φ₁))    &(\suc~n) &= \fuh~(\fstrip₁~φ₁~n)\\
  \fstrip₁ &(¬ ⊥)         &(\suc~n) &= ⊤\\
  \fstrip₁ &(¬ ⊤)         &(\suc~n) &= ⊥\\
  \fstrip₁ &φ             &n        &= φ.
  \end{array}
  \end{aligned}
\end{empheq}

\begin{mainlemma}
\label{lem:lem-inv-strip}
Let $n : \Nat$. If $Γ~⊢~\fstrip₁~φ~n$ then $Γ~⊢~φ$.
\end{mainlemma}

@@

\begin{mainth}
\label{thm:strip}
Let $s_1$ and $s_2$ be the subgoals of the goal $φ$, that is,
$\fstrip~\varphi≡~s_{1}∧~s_{2}$.
% $\fstrip₁~φ~n~≡~s₂∧~s₃$ where $\text{n} : \Nat$ is the complexity measure $\fstrip_{cm}$.
If $Γ ⊢ s₂$ and $Γ ⊢ s₃$ then $Γ ⊢ φ$.
\end{mainth}
% % No creo que sea necesario presentar las demostraciones
% \begin{proof}
% \begin{equation*}
%   \begin{bprooftree}
%   \AxiomC{$ Γ ⊢ s_1 $}
%   \AxiomC{$ Γ ⊢ s_2 $}
%   \RightLabel{∧-intro}
%   \BinaryInfC{$Γ ⊢ s_1\wedge s_2$}
%   \AxiomC{$\fstrip~φ ≡ s_1\wedge s_2 $}
%   \RightLabel{By def.~\eqref{eq:strip}}
%   \UnaryInfC{$\fstrip₁~φ~(\fstrip_{cm}~\varphi) ≡ s_1\wedge s_2 $}
%   \RightLabel{\fsubst}
%   \BinaryInfC{$Γ ⊢ \fstrip₁~φ~n$}
%   \RightLabel{Lemma~\ref{lem:lem-inv-strip}}
%   \UnaryInfC{$Γ ⊢ φ$}
% \end{bprooftree}
% \end{equation*}
% \end{proof}

@@

The user of \Metis will probably note another rule in the derivations related
with these subgoals, the \negate rule. Since \Metis proves a conjecture by
refutation, to prove a subgoal, \Metis assumes its negation using the \negate
rule. This rule always ocurrs in the \TSTP derivation after the application of
the \strip rule per each subgoal.

>>>>

\begin{notation}
$\Bound$ is a synonym for $\Nat$ type to refer to a natural number that constrains
the number of recursive calls for the function that use it in its type.
\end{notation}

\begin{mainlemma}
  \label{lem:inv-uh-lem}
Let be $n : \Bound$. If $Γ ⊢ \fuh₁~φ~n$ then $Γ ⊢ φ$ where

\begin{empheq}[box=\fcolorbox{bocolor}{bgcolor}]{equation}
  \label{eq:uh-structured}
  \begin{aligned}
  &\hspace{.495mm} \fuh_{1} : \Prop → \Bound → \Prop\\
  &\begin{array}{llll}
  \fuh_{1} &(φ₁ ⇒ (φ₂ ⇒ φ₃)) &(\suc~n) &= \fuh_{1}~((φ₁ ∧ φ₂) ⇒ φ₃)~n\\
  \fuh_{1} &(φ₁ ⇒ (φ₂ ∧ φ₃)) &(\suc~n) &= \fuh_{1}~(φ₁ ⇒ φ₂)~n ∧ \fuh_{1}~(φ₁ ⇒ φ₃)~n\\
  \fuh_{1} &φ &n &= φ.
  \end{array}
  \end{aligned}
\end{empheq}
\end{mainlemma}

To provide such a number given a formula
$\varphi$, we call the $\fuh_{cm}$ function defined in~\eqref{eq:uh-complexity}.
For the convenience in the following descriptions, we have the
function $\fuh : \Prop → \Prop$ for $\varphi \mapsto \fuh_{1}~φ~(\fuh_{cm}~φ)$.

\begin{empheq}[box=\fcolorbox{bocolor}{bgcolor}]{equation}
  \label{eq:uh-complexity}
  \begin{aligned}
    &\hspace{.495mm}\fuh_{cm} : \Prop → \Bound\\
    &\begin{array}{llll}
    \fuh_{cm}~(φ₁ ⇒ (φ₂ ⇒ φ₃)) &= \fuh_{cm}~φ₃ + 2\\
    \fuh_{cm}~(φ₁ ⇒ (φ₂ ∧ φ₃)) &= \fmax~(\fuh_{cm}~φ₂)~(\fuh_{cm}~φ₃) + 1\\
    \fuh_{cm}~φ                &= 0.
    \end{array}
  \end{aligned}
\end{empheq}

Now, we define the $\fstrip$ function in \eqref{eq:strip} in a similar way as we
did above for the $\fuh$ function. We have a definition of the $\fstrip₁$ function
in~\eqref{eq:strip-fixed} and a function to find its bound for its recursive calls
by $\fstrip_{cm}$ defined in~\cite{Prieto-Cubides2017a}.
% \vspace*{-5mm}
\begin{empheq}[box=\fcolorbox{bocolor}{bgcolor}]{align}
  \label{eq:strip}
  \begin{split}
  &\fstrip : \Prop → \Prop\\
  &\fstrip~φ~ = \fstrip₁~φ~(\fstrip_{cm}~φ).
  \end{split}
\end{empheq}

The $\fstrip$ function yields the conjunction of subgoals that implies the goal
of the problem in the \Metis \TSTP derivations. Its definition comes mainly from
the reading of the \Metis source code and some examples from \TSTP derivations.

\begin{empheq}[box=\fcolorbox{bocolor}{bgcolor}]{equation}
  \label{eq:strip-fixed}
  \begin{aligned}
  &\hspace{.495mm}\fstrip₁ : \Prop → \Bound → \Prop\\
  &\begin{array}{llll}
  \fstrip₁ &(φ₁ ∧ φ₂)     &(\suc~n) &= \fuh~(\fstrip₁~φ₁~n) ∧ \fuh~(φ₁ ⇒ \fstrip₁~φ₂~n)\\
  \fstrip₁ &(φ₁ ∨ φ₂)     &(\suc~n) &= \fuh~((¬ φ₁) ⇒ \fstrip₁~φ₂~n)\\
  \fstrip₁ &(φ₁ ⇒ φ₂)     &(\suc~n) &= \fuh~(φ₁ ⇒ \fstrip₁~φ₂~n)\\
  \fstrip₁ &(¬ (φ₁ ∧ φ₂)) &(\suc~n) &= \fuh~(φ₁ ⇒ \fstrip₁~(¬ φ₂)~n)\\
  \fstrip₁ &(¬ (φ₁ ∨ φ₂)) &(\suc~n) &= \fuh~(\fstrip₁~(¬ φ₁)~n) ∧ \fuh~((¬ φ₁) ⇒ \fstrip₁~(¬ φ₂)~n)\\
  \fstrip₁ &(¬ (φ₁ ⇒ φ₂)) &(\suc~n) &= \fuh~(\fstrip₁~φ₁~n) ∧ \fuh~(φ₁ ⇒ \fstrip₁~(¬ φ₂)~n)\\
  \fstrip₁ &(¬ (¬ φ₁))    &(\suc~n) &= \fuh~(\fstrip₁~φ₁~n)\\
  \fstrip₁ &(¬ ⊥)         &(\suc~n) &= ⊤\\
  \fstrip₁ &(¬ ⊤)         &(\suc~n) &= ⊥\\
  \fstrip₁ &φ             &n        &= φ.
  \end{array}
  \end{aligned}
\end{empheq}

\begin{mainlemma}
\label{lem:lem-inv-strip}
Let $n : \Bound$. If $Γ~⊢~\fstrip₁~φ~n$ then $Γ~⊢~φ$.
\end{mainlemma}

@@

\begin{mainth}
\label{thm:strip}
Let $s_1$ and $s_2$ be the subgoals of the goal $φ$, that is,
$\fstrip~\varphi≡~s_{1}∧~s_{2}$.
% $\fstrip₁~φ~n~≡~s₂∧~s₃$ where $\text{n} : \Nat$ is the complexity measure $\fstrip_{cm}$.
If $Γ ⊢ s_{1}$ and $Γ ⊢ s_{2}$ then $Γ ⊢ φ$.
\end{mainth}

% % No creo que sea necesario presentar las demostraciones
% \begin{proof}
% \begin{equation*}
%   \begin{bprooftree}
%   \AxiomC{$ Γ ⊢ s_1 $}
%   \AxiomC{$ Γ ⊢ s_2 $}
%   \RightLabel{∧-intro}
%   \BinaryInfC{$Γ ⊢ s_1\wedge s_2$}
%   \AxiomC{$\fstrip~φ ≡ s_1\wedge s_2 $}
%   \RightLabel{By def.~\eqref{eq:strip}}
%   \UnaryInfC{$\fstrip₁~φ~(\fstrip_{cm}~\varphi) ≡ s_1\wedge s_2 $}
%   \RightLabel{\fsubst}
%   \BinaryInfC{$Γ ⊢ \fstrip₁~φ~n$}
%   \RightLabel{Lemma~\ref{lem:lem-inv-strip}}
%   \UnaryInfC{$Γ ⊢ φ$}
% \end{bprooftree}
% \end{equation*}
% \end{proof}

@@

The user of \Metis will probably note another rule in the derivations related
with these subgoals, the \negate rule. Since \Metis proves a conjecture by
refutation, to prove a subgoal, \Metis assumes its negation using the \negate
rule. We will find this rule always after the \strip rule per each subgoal in the
\TSTP derivation.
